---
title: "Reproducible Research Assignment 1"
output: html_document
---

## Load and Preprocess the Data

```{r}

#load packages
library(ggplot2)
library(dplyr)
library(zoo)
library(data.table)
# Create a temporary directory to store the file

tf <- tempfile()
td <- tempdir()

# Load the file

zipData <- "https://github.com/JohnDuncanData/RepData_PeerAssessment1/raw/master/activity.zip"

download.file(zipData, tf, mode="wb")

#Unzip the zip file

FileNames <- unzip(tf, exdir=td)

#read the file as a table

FitData <- read.csv(FileNames[1])

knitr::opts_chunk$set(echo = TRUE)
```

## What is the mean total number of steps taken per day?

```{r}

#Remove NA values and summarise days

FitDatanoNA <- FitData[complete.cases(FitData), ]

SumFitData <- FitDatanoNA %>%
  group_by(date) %>%
  summarise(steps=sum(steps))

#create charts and summary statistics

TotalStepHist <- ggplot(SumFitData, aes(x=steps)) + geom_histogram(binwidth = 2500) +ggtitle("Histogram of Total Steps per Day")
TotalStepHist

TotalStepsMean <- mean(SumFitData$steps)
TotalStepsMean

TotalStepsMedian <- median(SumFitData$steps)
TotalStepsMedian

```

## What is the Average Daily Activity Pattern

```{r}
IntervalFitData <- FitDatanoNA %>%
  group_by(interval) %>%
  summarise(steps=mean(steps))

IntervalTimeSeries <- ggplot(IntervalFitData, aes(x=interval, y=steps)) + geom_line() +ggtitle("Average Steps at each Interval")+ylab("Average Steps")

IntervalFitDataRank<- IntervalFitData[order(-IntervalFitData$steps),]
head(IntervalFitDataRank)
TopInterval <- IntervalFitDataRank$interval[1]

IntervalTimeSeries
TopInterval

```

## Input Missing Values
```{r}

FitDataNA <- FitData[!complete.cases(FitData), ]
NACount <- length(FitDataNA$steps)
NACount

# Replace NA with overall mean steps

FitDataTB <- data.table(FitData)
FitDataFixed <- FitDataTB[, steps := na.aggregate(steps)]

SumFitDataFixed <- FitDataFixed %>%
  group_by(date) %>%
  summarise(steps=sum(steps))

#create charts and summary statistics

TotalStepHist <- ggplot(SumFitDataFixed, aes(x=steps)) + geom_histogram(binwidth = 2500) +ggtitle("Histogram of Total Steps per Day")
TotalStepHist

TotalStepsMean <- mean(SumFitDataFixed$steps)
TotalStepsMean

TotalStepsMedian <- median(SumFitDataFixed$steps)
TotalStepsMedian



```
We can see that replacing the NA values with the average has shifted the median to equal the mean 

## Are there differences in activity patterns on Weekdays and Weekends?

```{r}

FitDataFixed$date2 <- as.Date(FitDataFixed$date)
Weekday <- weekdays(FitDataFixed$date2)

FitDataFixed$Weekend <- with(FitDataFixed, ifelse(
  Weekday == "Saturday", "Weekend", ifelse(
    Weekday == "Sunday", "Weekend", "Weekday")
)
)

WeekendFitData <- FitDataFixed %>%
  group_by(interval,Weekend) %>%
  summarise(steps=mean(steps))

Weekendplot <- ggplot(WeekendFitData, aes(x=interval, y=steps)) +geom_line()+facet_grid(Weekend ~.)+ylab("Average Steps")+ggtitle("Average Steps by Interval - Weekend vs Weekday")
Weekendplot


